---
title: "RR Skills (3): Statistical Modeling "
author: Andrey Koval
date: November 18, 2014
output: 
  ioslides_presentation:
    widescreen: true
    transition: faster
---

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
#   fig.height = 4.8, 
#   fig.width = 6.5, 
  out.width = NULL,
  fig.path = 'figure_rmd_try/',     
  dev = "png",
  dpi = 70
)
echoChunks <- FALSE
warningChunks<- FALSE
messageChunks<- FALSE
outwidthChunks <- "90%"
options(width=120) #So the output is 50% wider than the default.
read_chunk("./Reports/2014-11-18-Statistical-Modeling.R") # the file to which knitr calls for the chunks
```


## Overview of the Series

- Oct 14  -- Intro to Reproducible Research  
- Oct 21  -- RR Basic Skills (1): Data Manipulation
- Oct 28  -- Intro to Latent Class and Latent Transition Models  
- Nov  4  -- RR Basic Skills (2): Graph Production  
- Nov 11      
- Nov 18  -- **RR Basic Skills (3): Statistical Modeling**   
- Nov 25  -- RR Basic Skills (4): Dynamic Reporting   
- Dec 2 -- Migrating into R from other Statistical Software  


## Previously:

<img src = 'images/ai/DataMap_full.png' width="100%"></img> 
```p
"./Scripts/Data/dsL.R"
```

download the files to work along at [GitHub](https://github.com/IALSA/COAG-colloquium-2014F)

## Load Data {.smaller}
```{r LoadPackages, echo=F, warning=F, message=F, results='hide'}
```

```{r LoadData, echo=T, warning=F, message=F, results='hide'}
```
```{r}
str(dsL)
```

## Load graphical theme {.smaller}
```{r loadTheme, out.width="95%", warning=F}
```


## Mapping data to graphics {.smaller}
<div class="columns-2">
```{r dsM01}
```
</br>
```{r graphBasic, out.width="95%", warning=F, echo=F}
```



</div>

<math xmlns="http://www.w3.org/1998/Math/MathML" display="block">
<math display='block' xmlns='http://www.w3.org/1998/Math/MathML'>
 <mrow>
  <msub>
   <mi>y</mi>
   <mrow>
    <mi>i</mi><mi>t</mi></mrow>
  </msub>
  </mrow>
</math>

Press(P): Note  
Next: final graph from last time 




## Variables and Aesthetics {.smaller}
<div class="columns-2">
```{r graphFullPrevious, out.width="95%", warning=F}
```
</div>
Press (P): Zoom  
Next: What is modeling
<div class="notes">
Last time we've studied the anatomy of ggplot and saw how developing our graph layer by layer we can achive a rather impressive results of data visualization in relatively short amout of time.
```{r graphFullPrevious, out.width="100%", echo=F,warning=F,message=F}
```
</div>

## Equation 








