---
title: "RR Skills (2): Graph Production "
author: Andrey Koval
date: November 4, 2014
output: 
  ioslides_presentation:
    widescreen: true
    transition: faster
    css: custom.css
---

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
#   fig.height = 4.8, 
#   fig.width = 6.5, 
  out.width = NULL,
  fig.path = 'figure_rmd_try/',     
  dev = "png",
  dpi = 400
)
echoChunks <- FALSE
warningChunks<- FALSE
messageChunks<- FALSE
outwidthChunks <- "90%"
options(width=120) #So the output is 50% wider than the default.
read_chunk("./Reports/2014-11-04-Producing-Graphs.R") # the file to which knitr calls for the chunks
```

<!--
## Load Data {.smaller}
```{r LoadData, echo=T, warning=F, message=F, results='hide'}
```
```{r}
str(dsL)
```


## Load Data
```{r}
dplyr::filter(dsL, id==1)
```

## Load Data
```{r}
dplyr::filter(dsL, id==1) %>% select(id, year, attend, attendF)
```


## Graph 0 {.smaller}
<div class="columns-2">
```{r dsM01, out.width="50%"}
```

```{r, out.width="95%", eval=FALSE}
p <- ggplot2::ggplot(dsM,aes(x=year,y=attend))
p
```
</br> </br> </br> </br> </br> </br> </br> </br> </br> </br>
```
# Error: No layers in plot
```
</div>


## Define theme {.smaller}
```{r loadTheme, echo=T, warning=F, message=F, out.width="60%"}
```

see more options for [themes in ggplot2](http://docs.ggplot2.org/dev/vignettes/themes.html)
-->   





## Faceting {.smaller}
<div class="columns-2">
```{r graph22, out.width="95%", warning=F}
```

</br> </br> </br>
Next: removing factor level  
Note: press "p" for zoom    
</div>
<div class="notes">
```{r graph22, out.width="100%", echo=F,warning=F,message=F}
```
</div>


## Exclude levels {.smaller}
<div class="columns-2">
```{r graph23, out.width="95%", warning=F}
```

</br> </br> </br>
Next: tranforming variable    
Note: press "p" for zoom    
</div>
</br> 
<div class="notes">
```{r graph23, out.width="100%", echo=F,warning=F,message=F}
```
</div>



## Centering time {.smaller}
<div class="columns-2">
```{r graph24, out.width="95%", warning=F}
```

</br> </br> </br> 
Next: final styling  
Note: press "p" for zoom  
</div>
<div class="notes">
```{r graph24, out.width="100%", echo=F,warning=F,message=F}
```
</div>



## Final touches {.smaller}
<div class="columns-2">
```{r graph25, out.width="95%", warning=F}
```

.
</br> </br> </br> </br>
Next: final graph  
</div>















## Question? Comments?

You can forward questions to Ann Greenwood (ann.greenwood@popdata.bc.ca)  or Vincenza Gruppuso (vincenza@uvic.ca) during the live Q&A. There are standing by to take your questions.   
 
